---
title: "Statistical rethinking assignements - Chapter 3"
author: "Deportivo Bayes"
date: "July 24, 2020"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#library(brms)
```

##Easy
These problems use the samples from the posterior distribution for the globe tossing example.
This code will give you a specific set of samples, so that you can check your answers exactly

```{r Easy}
length.out <- 1001
n_water <- 6
n_trials  <- 9

d <-
  tibble(p_grid     = seq(from = 0, to = 1, length.out = length.out),
         prior      = 1) %>% 
  mutate(likelihood = dbinom(n_water, size = n_trials, prob = p_grid)) %>% 
  mutate(posterior  = (likelihood * prior) / sum(likelihood * prior))

# How many samples would you like?
n_samples <- 1e4

# Make it reproducible
set.seed(3)

# Draw samples
samples <-
  d %>% 
  sample_n(size = n_samples, weight = posterior, replace = T)
```

**3E1.** How much posterior probability lies below p = 0.2?

**SOLUTION**

```{r 3E1}
solution <- 
  samples %>% 
  filter(p_grid < .2) %>% 
  summarise(sum = n() / n_samples)

print(paste0("The posterior probability that lies below p=0.2 is ", solution$sum))
```

**3E2.** How much posterior probability lies above p = 0.8?

**SOLUTION**

```{r 3E2}
solution <- 
  samples %>% 
  filter(p_grid > .8) %>% 
  summarise(sum = n() / n_samples)

print(paste0("The posterior probability that lies above p=0.8 is ", solution$sum))
```

**3E3.** How much posterior probability lies between p = 0.2 and p = 0.8?

**SOLUTION**

```{r 3E3}
solution <- 
  samples %>% 
  filter(p_grid > .2 & p_grid < .8) %>% 
  summarise(sum = n() / n_samples)

print(paste0("The posterior probability that lies above p=0.8 is ", solution$sum))
```

**3E4.** 20% of the posterior probability lies below which value of p?

**SOLUTION**

```{r 3E4}
solution <- 
  samples %>% 
  filter(p_grid > .2 & p_grid < .8) %>% 
  summarise(sum = n() / n_samples)

print(paste0("The posterior probability that lies above p=0.8 is ", solution$sum))
```