---
title: "Statistical rethinking assignements - Chapter 3"
author: "Deportivo Bayes"
date: "July 24, 2020"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rethinking)
```

##Easy
These problems use the samples from the posterior distribution for the globe tossing example.
This code will give you a specific set of samples, so that you can check your answers exactly

```{r Easy}
length.out <- 1001
n_water <- 6
n_trials  <- 9

d <-
  tibble(p_grid     = seq(from = 0, to = 1, length.out = length.out),
         prior      = 1) %>% 
  mutate(likelihood = dbinom(n_water, size = n_trials, prob = p_grid)) %>% 
  mutate(posterior  = (likelihood * prior) / sum(likelihood * prior))

# How many samples would you like?
n_samples <- 1e4

# Make it reproducible
set.seed(3)

# Draw samples
samples <-
  d %>% 
  sample_n(size = n_samples, weight = posterior, replace = T)
```

**3E1.** How much posterior probability lies below p = 0.2?

**SOLUTION**

```{r 3E1}
solution <- 
  samples %>% 
  filter(p_grid < .2) %>% 
  summarise(sum = n() / n_samples)

print(paste0("The posterior probability that lies below p=0.2 is ", solution$sum))
```

**3E2.** How much posterior probability lies above p = 0.8?

**SOLUTION**

```{r 3E2}
solution <- 
  samples %>% 
  filter(p_grid > .8) %>% 
  summarise(sum = n() / n_samples)

print(paste0("The posterior probability that lies above p=0.8 is ", solution$sum))
```

**3E3.** How much posterior probability lies between p = 0.2 and p = 0.8?

**SOLUTION**

```{r 3E3}
solution <- 
  samples %>% 
  filter(p_grid > .2 & p_grid < .8) %>% 
  summarise(sum = n() / n_samples)

print(paste0("The posterior probability that lies above p=0.8 is ", solution$sum))
```

**3E4.** 20% of the posterior probability lies below which value of p?

**SOLUTION**

```{r 3E4}
solution <- 
  samples %>% 
  filter(p_grid > .2 & p_grid < .8) %>% 
  summarise(sum = n() / n_samples)

print(paste0("The posterior probability that lies above p=0.8 is ", solution$sum))
```

**3E5.** 20% of the posterior probability lies above which value of p?

**SOLUTION**
```{r 3E5}
solution <- 
  samples %>% 
  filter(p_grid > .2 & p_grid < .8) %>% 
  summarise(sum = n() / n_samples)

print(paste0("The posterior probability that lies above p=0.8 is ", solution$sum))
```

**3E6.** Which values of p contain the narrowest interval equal to 66% of the posterior probability?

**SOLUTION**
```{r 3E6}
solution <- 
  samples %>% 
  select(p_grid) %>% 
  pull() %>% 
  HPDI(prob = 0.66)

print(paste0("The values of p between ", round(solution[1],2), " and ", round(solution[2],2), " contain the narrowest interval equal to 66% of the posterior probability "))
```

**3E7.** Which values of p contain 66% of the posterior probability, assuming equal posterior probability
both below and above the interval?

**SOLUTION**
```{r 3E7}
solution <- 
  samples %>% 
  select(p_grid) %>% 
  pull() %>% 
  PI(prob = 0.66)

print(paste0("The values of p between ", round(solution[1],2), " and ", round(solution[2],2), " contain 66% of the posterior probability "))
```

##Medium
**3M1.** Suppose the globe tossing data had turned out to be 8 water in 15 tosses. Construct the posterior
distribution, using grid approximation. Use the same flat prior as before.

**SOLUTION**
```{r 3M1, fig.height = 4, fig.width = 6, fig.align = "left"}
length.out <- 1001
n_water <- 8
n_trials  <- 15

data3M1 <-
  tibble(p_grid     = seq(from = 0, to = 1, length.out = length.out),
         prior      = 1) %>% 
  mutate(likelihood = dbinom(n_water, size = n_trials, prob = p_grid)) %>% 
  mutate(posterior  = (likelihood * prior) / sum(likelihood * prior))

ggplot(data = data3M1, 
       aes(x = p_grid, y = posterior)) +
  geom_line() +
  labs(x = "Proportion water",
       y = "Density") +
  theme_minimal() + 
  ggtitle("Poteriors with 8 waters in 15 tosses")
```

**3M2.** Draw 10,000 samples from the grid approximation from above. Then use the samples to calculate
the 90% HPDI for p.

**SOLUTION**

We will first calculate the HPDI
```{r 3M2.1}
# How many samples would you like?
n_samples <- 1e4

# Make it reproducible
set.seed(3)

# Draw samples
samples3M2 <-
  data3M1 %>% 
  sample_n(size = n_samples, weight = posterior, replace = T)

solution <- 
  samples3M2 %>% 
  select(p_grid) %>% 
  pull() %>% 
  HPDI(prob = 0.90)

print(paste0("The values of p between ", round(solution[1],2), " and ", round(solution[2],2), " contain the narrowest interval equal to 66% of the posterior probability "))
```

And now we will plot the interval in along with the sampled distribution.

```{r 3M2.2, fig.height = 4, fig.width = 6, fig.align = "left"}
ggplot(data = samples3M2, 
       aes(x = p_grid)) +
  geom_density(fill = "red", 
               alpha = .3) +
  geom_ribbon(data = samples3M2 %>% filter(p_grid < solution[2] & p_grid > solution[1]),
              aes(ymin = 0, ymax = posterior*n_samples/10),
              fill = "red", 
              alpha = .7) +
  geom_line(data = data3M1, 
            aes(x = p_grid, y = posterior*n_samples/10), 
            linetype = "dashed") +
  labs(x = "Proportion water (p)",
       y = "Density") +
  theme_minimal() + 
  ggtitle("HPDI of the sample distribution with 8 waters in 15 tosses")
```

**3M3.** Construct a posterior predictive check for this model and data. This means simulate the distribution
of samples, averaging over the posterior uncertainty in p. What is the probability of observing
8 water in 15 tosses?

**SOLUTION**

Let's plot again the posterior of **3M1.**:

```{r 3M3.1, fig.height = 4, fig.width = 6, fig.align = "left"}
ggplot(data = data3M1, 
       aes(x = p_grid, y = posterior)) +
  geom_line() +
  labs(x = "Proportion water",
       y = "Density") +
    geom_segment(data = . %>% 
                 filter(p_grid %in% c(seq(from = .1, to = .9, by = .1), 3 / 10)),
                 aes(xend = p_grid,
                   y = 0, yend = posterior),
               size = .3, linetype = "dashed", 
               color = "black", show.legend = F) +
  theme_minimal() + 
  ggtitle("Posterior with 8 waters in 15 tosses")
```

Now we have the posterior we should draw samples with each of the ten possible parameter values.

```{r 3M3.2, fig.height = 4, fig.width = 6, fig.align = "left"}
# Number of draws for each parameter value
n_draws <- 1e5

# Make it replicable
set.seed(3)

# A function to extract the number of waters in the draw
simulate_binom <- function(probability){
  rbinom(n_draws, size = 9, prob = probability) 
}

# A tibble with the number of waters for each of the parameter values
data_parameter <-
  tibble(probability = seq(from = .1, to = .9, by = .1)) %>% 
  mutate(draws       = purrr::map(probability, simulate_binom)) %>% 
  unnest(draws) %>% 
  mutate(label       = str_c("p = ", probability))

head(data_parameter)
```

and now let's plot the samples for each parameter value.

```{r 3M3.3, fig.height = 2.5, fig.width = 10, fig.align = "left"}
data_parameter %>%
  ggplot(aes(x = draws)) +
  geom_histogram(binwidth = 1, center = 0,
                 color = "grey92", size = 1/10) +
  scale_x_continuous(NULL, breaks = seq(from = 0, to = 9, by = 3)) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(subtitle = "Sampling distributions") +
  coord_cartesian(xlim = c(0,9)) +
  theme_minimal() +
  facet_wrap(~ label, ncol = 9) 
```

But... Why should we use only ten possible parameter values if our grid space has a thousand steps? Let's repeat the last step but using a thousan possible values, that is, approximating better the integral form of the Posterior predictive distribution: $$p(\hat{x}|data) = \int_{\Theta} p(\hat{x}|\Theta,data) p(\Theta|data) d\Theta$$

Whrere $\hat{x}$ is the sampled data, $\Theta$ the estimated parameter (percentage of water in our case) and $p(\Theta|data)$ the posterior distribution.

**3M4.** Using the posterior distribution constructed from the new (8/15) data, now calculate the probability
of observing 6 water in 9 tosses.

**SOLUTION**

**3M5.** Start over at 3M1, but now use a prior that is zero below p = 0.5 and a constant above p = 0.5.
This corresponds to prior information that a majority of the Earth’s surface is water. Repeat each
problem above and compare the inferences. What difference does the better prior make? If it helps,
compare inferences (using both priors) to the true value p = 0.7.

**SOLUTION**